---
title: "Storage Capacity Analysis"
author: "Nicolas Genibre"
date: "September 26, 2016"
output:
  word_document: default
  pdf_document:
    fig_caption: yes
    fig_height: 4
    fig_width: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

What is the Storage Capacity looks like for the NetApp cluster in Fort Lauderdale Laboratory?

For this Analysis we implemented the NetApp Perfstat tool to run on a schedule, for 16 days, and captured 48 iterations of 15 minutes each day using the following to determine capacity:

df -x -m (pre and prostats); df -A -m (pre and prostats); sysstat_x_1sec ; sysstat_x_5sec

## Clean Data

The raw data is composed of 32 files, 16 files per NetApp controller. Each files are plain text output containing the iterations for a day.  

We're cleaning the data with the help of shell scripts to transform the data into a csv type file:

aggr.sh; vol.sh and the result is in the data directory as aggr.csv and vol.csv

```{r}
myaggr <- read.csv("../data/aggr.csv")

myvol <- read.csv("../data/vol.csv")
```

The **Time** is expressed in epoch date. There are two **Filers** in the Cluster.

**Total**, **Used** and **Free** capacity are measurements in MB. The **Ratio** is the expression in percentage of Used versus Total.


## Aggregate

```{r message=FALSE}

library(dplyr)
library(ggplot2)

```

```{r }

head(myaggr)

```

```{r echo=FALSE}

myaggr$Total <- myaggr$Total / 1024
myaggr$Used <- myaggr$Used / 1024
myaggr$Time <- as.POSIXct(myaggr$Time, origin="1970-01-01")

myaggrf1 <- filter(myaggr, Filer == "netapp-01")
myaggrf2 <- filter(myaggr, Filer == "netapp-02")

myaggr0f1 <- filter(myaggr, Aggregate == "n1_aggr0")
myaggr1f1 <- filter(myaggr, Aggregate == "n1_aggr1")
myaggr2f1 <- filter(myaggr, Aggregate == "n1_aggr2")
myaggr0f2 <- filter(myaggr, Aggregate == "n2_aggr0")
myaggr1f2 <- filter(myaggr, Aggregate == "n2_aggr1")

```

```{r echo=FALSE}

p <- ggplot(myaggr, aes(Time)) + geom_line(aes(y = Used, colour = Aggregate)) + geom_line(aes(y = mean(myaggr1f1$Total), colour = "Total n1_aggr1")) + geom_line(aes(y = mean(myaggr2f1$Total), colour = "Total n1_aggr2")) + geom_line(aes(y = mean(myaggr1f2$Total), colour = "Total n2_aggr1")) 

p + labs(title = "Aggregate Usage") + labs(x = "Time", y = "Size (GB)") + labs(colour="") + scale_color_manual(values=c("n1_aggr0"="#99d8c9", "n1_aggr1"="#de2d26", "n1_aggr2"="#d95f0e", "n2_aggr0"="#99d8c9", "n2_aggr1"="#756bb1", "Total n1_aggr1"="#de2d26", "Total n1_aggr2"="#d95f0e", "Total n2_aggr1"="#756bb1"))

```

The Aggregate' used capacity has not change much from the beginning to the end of the data collection.

According the linear model we see a slight decrease of Aggregate 1 and increase of Aggregate 2 on Filer 1 over time. On Filer 2 the increase is larger for the Arggregate 2.

```{r echo=FALSE}

g <- ggplot(myaggr, aes(Time, Used))

#g + geom_line(aes(colour = Aggregate)) + labs(title = "Aggregate Usage") + labs(x = "Time", y = "Size (GB)") + facet_grid( . ~ Filer)

g + geom_smooth(aes(colour = Aggregate), method = "lm") + labs(title = "Aggregate Linear Model Usage") + labs(x = "Time", y = "Size (GB)") + facet_grid( . ~ Filer)

g <- ggplot(myaggr, aes(Time, UsageRatio))

g + geom_line(aes(colour = Aggregate)) + labs(title = "Usage Ratio") + labs(x = "Time", y = "%") + facet_grid( . ~ Filer)
```

Aggregates are kept under the 85% threshold. During the collection we can see that one of the Aggregate on Filer 1 was increased.


## Volume



```{r echo=FALSE}

myvol$Total <- myvol$Total / 1024
#myvol$Used <- as.numeric(as.character(myvol$Used)) 
myvol$Used <- myvol$Used / 1024
myvol$Time <- as.POSIXct(myvol$Time, origin="1970-01-01")
myvolf1 <- filter(myvol, Filer == "netapp-01")
myvolf2 <- filter(myvol, Filer == "netapp-02")

```


```{r echo=FALSE}

g <- ggplot(myvolf1, aes(Time, Used))

g + geom_line(aes(colour = Volume)) + labs(title = "Volume Usage") + labs(x = "Time", y = "Size (GB)") + facet_grid( . ~ Filer) + theme(legend.position="bottom") + theme(legend.title=element_blank()) + theme(legend.text = element_text(size=5))

g <- ggplot(myvolf2, aes(Time, Used))

g + geom_line(aes(colour = Volume)) + labs(title = "Volume Usage") + labs(x = "Time", y = "Size (GB)")  + facet_grid( . ~ Filer) + theme(legend.position="bottom") + theme(legend.title=element_blank()) + theme(legend.text = element_text(size=7))


```

The top 5 volumes using the most capacity, and that are the most active, are the Cumulus volume, as we expected.

```{r echo=FALSE}

g <- ggplot(myvolf1, aes(Time, Used))


g + geom_smooth(aes(colour = Volume), method = "lm") + labs(title = "Volume Linear Model Usage") + labs(x = "Time", y = "Size (GB)") + facet_grid( . ~ Filer) + theme(legend.position="bottom") + theme(legend.title=element_blank()) + theme(legend.text = element_text(size=5))

g <- ggplot(myvolf2, aes(Time, Used))


g + geom_smooth(aes(colour = Volume), method = "lm") + labs(title = "Volume Linear Model Usage") + labs(x = "Time", y = "Size (GB)") + facet_grid( . ~ Filer) + theme(legend.position="bottom") + theme(legend.title=element_blank()) + theme(legend.text = element_text(size=7))


```

According the Model one volume, vs1_cumulus_clus3, has shown the biggest increase during its activity 

```{r echo=FALSE}

g <- ggplot(myvol, aes(Time, UsageRatio))

g + geom_line(aes(colour = Volume)) + labs(title = "Usage Ratio") + labs(x = "Time", y = "%") + theme(legend.position='none')

```

The volumes vs1_labvdi2, vs1_ftlvmw1, vs1_xspool3, vs1_ftlvm2_vol are above the 85% threshold and that happens during the all period of the data collection.

```{r echo=FALSE}

#myvol$Time <- as.Date(as.character(myvol$Time), "%Y-%m-%d")

#head(myvol)

#voldaymean <- aggregate(Used ~ Time, myvol, mean)

#hist(voldaymean$Used, main="Average Space Used per Day", xlab="Size (GB) / Day", col="blue")


```